---
title: "FMQ Cancer"
author: "Workforce and Pay Analytics"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: word_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```


```{r, include=TRUE, echo=FALSE}
### 0 - Setup

## Run setup scripts where needed data and libraries are loaded
## output destination folder also created there

source(here::here("code", "00_setup.R"))

```


```{r, include=TRUE, echo=FALSE}
### 1 - Read in rds files that were prepared in setup file

consultant_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_consultant_wte.rds")),
  )
```


```{r, include=TRUE, echo=FALSE}
# specify and format the briefing rounds
new_quarter <- as.Date(paste0(briefing_round, "-01"), format = "%Y-%B-%d")
new_quarter <-format(new_quarter, format = "%B %Y")


prev_quarter <- as.Date(paste0(prevquarter_briefing_round, "-01"), format = "%Y-%B-%d")
prev_quarter <-format(prev_quarter, format = "%B %Y")
```

```{r, include=TRUE, echo=FALSE}

### For this FMQ, we include figures for oncology and radiology specialties

# Add in 2006 figure for Oncology consultants
oncology_cons_06 <- 69.542

# summarise data
# for previous quarter
oncology_cons_prev <- consultant_wte %>%
  filter(subjob_family %in% c("Clinical oncology", "Medical oncology")) %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

#for current quarter
oncology_cons_new <- consultant_wte %>%
  filter(subjob_family %in% c("Clinical oncology", "Medical oncology")) %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
oncology_prev_ndiff <- oncology_cons_prev - oncology_cons_06
oncology_new_ndiff <- oncology_cons_new - oncology_cons_06

# calculate the percentage difference with 2006
oncology_prev_pdiff <- (oncology_prev_ndiff/oncology_cons_06) *100
oncology_new_pdiff <- (oncology_new_ndiff/oncology_cons_06) *100



### Add in 2006 figure for Radiology consultants
radiology_cons_06 <- 223.525

# summarise data
# for previous quarter
radiology_cons_prev <- consultant_wte %>%
  filter(subjob_family == "Clinical radiology") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

#for current quarter
radiology_cons_new <- consultant_wte %>%
  filter(subjob_family == "Clinical radiology") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
radiology_prev_ndiff <- radiology_cons_prev - radiology_cons_06
radiology_new_ndiff <- radiology_cons_new - radiology_cons_06

# calculate the percentage difference with 2006
radiology_prev_pdiff <- (radiology_prev_ndiff/radiology_cons_06) *100
radiology_new_pdiff <- (radiology_new_ndiff/radiology_cons_06) *100
```


**Previous quarter (`r prev_quarter`):**

* Under the SNP, there has been a `r round_half_up(oncology_prev_pdiff, 1)`% increase in consultant oncologists (from `r round_half_up(oncology_cons_06, 1)` WTE in September 2006 to `r round_half_up(oncology_cons_prev, 1)` WTE in `r prev_quarter`).
* We've increased consultant radiologists by `r round_half_up(radiology_prev_pdiff, 1)`% (from `r round_half_up(radiology_cons_06, 1)` WTE in September 2006 to `r round_half_up(radiology_cons_prev, 1)` WTE in `r prev_quarter`).


**New quarter (`r new_quarter`):**

* Under the SNP, there has been a `r round_half_up(oncology_new_pdiff, 1)`% increase in consultant oncologists (from `r round_half_up(oncology_cons_06, 1)` WTE in September 2006 to `r round_half_up(oncology_cons_new, 1)` WTE in `r new_quarter`).
* We've increased consultant radiologists by `r round_half_up(radiology_new_pdiff, 1)`% (from `r round_half_up(radiology_cons_06, 1)` WTE in September 2006 to `r round_half_up(radiology_cons_new, 1)` WTE in `r new_quarter`).
