---
title: "FMQ Scottish Ambulance Service"
author: "Workforce and Pay Analytics"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```

```{r, include=TRUE, echo=FALSE}
### 0 - Setup

## Run setup scripts where needed data and libraries are loaded
## output destination folder also created there

source(here::here("code", "00_setup.R"))

```


```{r, include=TRUE, echo=FALSE}
### 1 - Read in rds files that were prepared in setup file

sas_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_sas_wte.rds")),
  )

# Load in wte and headcount by job family and board. This can be used for 2006 data
hc_wte_data <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_hc_wte_data.rds")),
  )
```


```{r, include=TRUE, echo=FALSE}

### For this FMQ, we include figures for staff in post in SAS, Paramedics and 
# ambulance technicians

# For SAS 
# Add in 2006 figure for SAS staff
sas_wte_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scottish Ambulance Service") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# For previous quarter
sas_wte_prevquarter <- sas_wte %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
sas_wte_new <- sas_wte %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
saswte_prev_ndiff <- sas_wte_prevquarter - sas_wte_06
saswte_new_ndiff <- sas_wte_new - sas_wte_06

# calculate the percentage difference with 2006
saswte_prev_pdiff <- (saswte_prev_ndiff/sas_wte_06) *100
saswte_new_pdiff <- (saswte_new_ndiff/sas_wte_06) *100


# For Paramedics
# For 2007 figure
paramedics_07 <- sas_wte %>%
  filter(subjob_family == "Paramedics") %>%
  summarise(sep07_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep07_sum)

# For previous quarter
paramedics_prevquarter <- sas_wte %>%
  filter(subjob_family == "Paramedics") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
paramedics_new <- sas_wte %>%
  filter(subjob_family == "Paramedics") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
paramedics_prev_ndiff <- paramedics_prevquarter - paramedics_07
paramedics_new_ndiff <- paramedics_new - paramedics_07

# calculate the percentage difference with 2007
paramedics_prev_pdiff <- (paramedics_prev_ndiff/paramedics_07) *100
paramedics_new_pdiff <- (paramedics_new_ndiff/paramedics_07) *100


# For Ambulance Technicians
# For 2007 figure
ambulancetech_07 <- sas_wte %>%
  filter(subjob_family == "Technician") %>%
  summarise(sep07_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep07_sum)

# For previous quarter
ambulancetech_prevquarter <- sas_wte %>%
  filter(subjob_family == "Technician") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
ambulancetech_new <- sas_wte %>%
  filter(subjob_family == "Technician") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
ambulancetech_prev_ndiff <- ambulancetech_prevquarter - ambulancetech_07
ambulancetech_new_ndiff <- ambulancetech_new - ambulancetech_07

# calculate the percentage difference with 2007
ambulancetech_prev_pdiff <- (ambulancetech_prev_ndiff/ambulancetech_07) *100
ambulancetech_new_pdiff <- (ambulancetech_new_ndiff/ambulancetech_07) *100
```


**Previous quarter (`r prev_quarter`):**  
*	Scottish Ambulance Service staffing is `r ifelse((saswte_prev_pdiff < 0), "down", "up")` `r round_half_up(saswte_prev_pdiff,1)`% under the SNP, by `r format(round_half_up(saswte_prev_ndiff, 1), big.mark =",")`  WTE between September 2006 and `r prev_quarter`, and now stands at `r format(round_half_up(sas_wte_prevquarter, 1), big.mark =",")` WTE.  
*	Paramedics are `r ifelse((paramedics_prev_pdiff < 0), "down", "up")` by `r round_half_up(paramedics_prev_pdiff,1)`% (`r round_half_up(paramedics_prev_ndiff,1)` WTE) from Sep 2007 to `r prev_quarter`, and now stand at `r format(round_half_up(paramedics_prevquarter, 1), big.mark =",")` WTE.    
*	Ambulance Technicians are `r ifelse((ambulancetech_prev_pdiff < 0), "down", "up")` by `r round_half_up(ambulancetech_prev_pdiff,1)`% (`r round_half_up(ambulancetech_prev_ndiff,1)` from Sep 2007 to `r prev_quarter`, and now stand at `r format(round_half_up(ambulancetech_prevquarter, 1), big.mark =",")` WTE.  

**New quarter (`r new_quarter`):**   
*	Scottish Ambulance Service staffing is `r ifelse((saswte_new_pdiff < 0), "down", "up")` `r round_half_up(saswte_new_pdiff,1)`% under the SNP, by `r format(round_half_up(saswte_new_ndiff, 1), big.mark =",")`  WTE between September 2006 and `r new_quarter`, and now stands at `r format(round_half_up(sas_wte_new, 1), big.mark =",")` WTE.  
*	Paramedics are `r ifelse((paramedics_new_pdiff < 0), "down", "up")` by `r round_half_up(paramedics_new_pdiff,1)`% (`r round_half_up(paramedics_new_ndiff,1)` WTE) from Sep 2007 to `r new_quarter`, and now stand at `r format(round_half_up(paramedics_new, 1), big.mark =",")` WTE.    
*	Ambulance Technicians are `r ifelse((ambulancetech_new_pdiff < 0), "down", "up")` by `r round_half_up(ambulancetech_new_pdiff,1)`% (`r round_half_up(ambulancetech_new_ndiff,1)` from Sep 2007 to `r new_quarter`, and now stand at `r format(round_half_up(ambulancetech_new, 1), big.mark =",")` WTE.  