---
title: "FMQ Maternity"
author: "Workforce and Pay Analytics"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```


```{r, include=TRUE, echo=FALSE}
### 0 - Setup

## Run setup scripts where needed data and libraries are loaded
## output destination folder also created there

source(here::here("code", "00_setup.R"))
```


```{r, include=TRUE, echo=FALSE}
### 1 - Read in rds files that were prepared in setup file

consultant_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_consultant_wte.rds")),
  )

nm_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_nm_wte.rds")),
  )
```


```{r, include=TRUE, echo=FALSE}

### For this FMQ, we include figures for qualified midwives and Obstetrics & Gynaecology consultants

# Add in 2006 figure for qualified midwives
#qual_midwives_06 <- 2798.962 #I can't verify this figure. Took it from old Scotland table

# summarise data
# for 2007
qual_midwives_07 <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

# for previous year
qual_midwives_prevyear <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# for previous quarter
qual_midwives_prevquarter <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

#for current quarter
qual_midwives_new <- nm_wte %>%
 filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
qualmidwives_prev_ndiff <- qual_midwives_prevquarter - qual_midwives_07
qualmidwives_new_ndiff <- qual_midwives_new - qual_midwives_07

# calculate the percentage difference with 2007
qualmidwives_prev_pdiff <- (qualmidwives_prev_ndiff/qual_midwives_07) *100
qualmidwives_new_pdiff <- (qualmidwives_new_ndiff/qual_midwives_07) *100

# calculate the difference and percentage with previous year
qualmidwives_prevyear_ndiff <- qual_midwives_new - qual_midwives_prevyear
qualmidwives_prevyear_pdiff <- (qualmidwives_prevyear_ndiff/qual_midwives_prevyear) *100

# calculate the difference and percentage with previous quarter
qualmidwives_prevqtr_ndiff <- qual_midwives_new - qual_midwives_prevquarter
qualmidwives_prevqtr_pdiff <- (qualmidwives_prevqtr_ndiff/qual_midwives_prevquarter) *100



### Add in 2006 figure for Obstetrics and Gynaecology consultants
obgyn_cons_06 <- 174.849

# summarise data

# for previous year
obgyn_cons_prevyear <- consultant_wte %>%
  filter(subjob_family == "Obstetrics & gynaecology") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# for previous quarter
obgyn_cons_prevquarter <- consultant_wte %>%
  filter(subjob_family == "Obstetrics & gynaecology") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

#for current quarter
obgyn_cons_new <- consultant_wte %>%
  filter(subjob_family == "Obstetrics & gynaecology") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
obgyn_prev_ndiff <- obgyn_cons_prevquarter - obgyn_cons_06
obgyn_new_ndiff <- obgyn_cons_new - obgyn_cons_06

# calculate the percentage difference with 2006
obgyn_prev_pdiff <- (obgyn_prev_ndiff/obgyn_cons_06) *100
obgyn_new_pdiff <- (obgyn_new_ndiff/obgyn_cons_06) *100

# calculate the difference and percentage with previous year
obgyn_prevyear_ndiff <- obgyn_cons_new - obgyn_cons_prevyear
obgyn_prevyear_pdiff <- (obgyn_prevyear_ndiff/obgyn_cons_prevyear) *100

# calculate the difference and percentage with previous quarter
obgyn_prevqtr_ndiff <- obgyn_cons_new - obgyn_cons_prevquarter
obgyn_prevqtr_pdiff <- (obgyn_prevqtr_ndiff/obgyn_cons_prevquarter) *100
```


**Previous quarter (`r prev_quarter`):**

* The number of qualified midwives as at `r prev_quarter` is `r format(round_half_up(qual_midwives_prevquarter, 1), big.mark = ",")` WTE. This is `r ifelse((qualmidwives_prev_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(qualmidwives_prev_ndiff, 1)` WTE (`r round_half_up(qualmidwives_prev_pdiff, 1)`%) from September 2007.
* As at `r prev_quarter`, there are `r round_half_up(obgyn_cons_prevquarter, 1)` WTE consultants in Obstetrics & Gynaecology, `r ifelse((obgyn_prev_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(obgyn_prev_ndiff, 1)` WTE (`r round_half_up(obgyn_prev_pdiff, 1)`%) from September 2006.


**New quarter (`r new_quarter`):**

* The number of qualified midwives as at `r new_quarter` is `r format(round_half_up(qual_midwives_new, 1), big.mark = ",")` WTE. This is `r ifelse((qualmidwives_new_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(qualmidwives_new_ndiff, 1)` WTE (`r round_half_up(qualmidwives_new_pdiff, 1)`%) from September 2007, `r ifelse((qualmidwives_prevyear_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(qualmidwives_prevyear_ndiff, 1)` WTE (`r round_half_up(qualmidwives_prevyear_pdiff, 1)`%) from `r prev_year` and `r ifelse((qualmidwives_prevqtr_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(qualmidwives_prevqtr_ndiff, 1)` WTE (`r round_half_up(qualmidwives_prevqtr_pdiff, 1)`%) from `r prev_quarter`.

* As at `r new_quarter`, there are `r round_half_up(obgyn_cons_new, 1)` WTE consultants in Obstetrics & Gynaecology, `r ifelse((obgyn_new_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(obgyn_new_ndiff, 1)` WTE (`r round_half_up(obgyn_new_pdiff, 1)`%) from September 2006, `r ifelse((obgyn_prevyear_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(obgyn_prevyear_ndiff, 1)` WTE (`r round_half_up(obgyn_prevyear_pdiff, 1)`%) from `r prev_year` and `r ifelse((obgyn_prevqtr_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(obgyn_prevqtr_ndiff, 1)` WTE (`r round_half_up(obgyn_prevqtr_pdiff, 1)`%) from `r prev_quarter`.