---
title: "FMQ State Hospital"
author: "Workforce and Pay Analytics"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```

```{r, include=TRUE, echo=FALSE}
### 0 - Setup

## Run setup scripts where needed data and libraries are loaded
## output destination folder also created there

source(here::here("code", "00_setup.R"))

```


```{r, include=TRUE, echo=FALSE}
### 1 - Read in rds files that were prepared in setup file

statehospital_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_statehospital_wte.rds")),
  )
```


```{r, include=TRUE, echo=FALSE}

### For this FMQ, we include figures for all staff in post in State Hospital, M&D, N&M and AHP and other staff in State Hospital

### For all State Hospital staff
# For 2007 figure
statehospital_07 <- statehospital_wte %>%
  summarise(sep07_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep07_sum)

# For previous quarter
statehospital_prevquarter <- statehospital_wte %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
statehospital_new <- statehospital_wte %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
statehospital_prev_ndiff <- statehospital_prevquarter - statehospital_07
statehospital_new_ndiff <- statehospital_new - statehospital_07

# calculate the percentage difference with 2006
statehospital_prev_pdiff <- (statehospital_prev_ndiff/statehospital_07) *100
statehospital_new_pdiff <- (statehospital_new_ndiff/statehospital_07) *100


### For M&D State Hospital staff
# For 2007 figure
md_statehospital_07 <- statehospital_wte %>%
  filter(job_family == "Medical & dental") %>%
  summarise(sep07_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep07_sum)

# For previous quarter
md_statehospital_prevquarter <- statehospital_wte %>%
  filter(job_family == "Medical & dental") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
md_statehospital_new <- statehospital_wte %>%
  filter(job_family == "Medical & dental") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
md_statehospital_prev_ndiff <- md_statehospital_prevquarter - md_statehospital_07
md_statehospital_new_ndiff <- md_statehospital_new - md_statehospital_07

# calculate the percentage difference with 2006
md_statehospital_prev_pdiff <- (md_statehospital_prev_ndiff/md_statehospital_07) *100
md_statehospital_new_pdiff <- (md_statehospital_new_ndiff/md_statehospital_07) *100


### For N&M State Hospital staff
# For 2007 figure
nm_statehospital_07 <- statehospital_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  summarise(sep07_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep07_sum)

# For previous quarter
nm_statehospital_prevquarter <- statehospital_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
nm_statehospital_new <- statehospital_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
nm_statehospital_prev_ndiff <- nm_statehospital_prevquarter - nm_statehospital_07
nm_statehospital_new_ndiff <- nm_statehospital_new - nm_statehospital_07

# calculate the percentage difference with 2006
nm_statehospital_prev_pdiff <- (nm_statehospital_prev_ndiff/nm_statehospital_07) *100
nm_statehospital_new_pdiff <- (nm_statehospital_new_ndiff/nm_statehospital_07) *100

### For AHP State Hospital staff
# For 2007 figure
ahp_statehospital_07 <- statehospital_wte %>%
  filter(job_family == "Allied health professions") %>%
  summarise(sep07_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep07_sum)

# For previous quarter
ahp_statehospital_prevquarter <- statehospital_wte %>%
  filter(job_family == "Allied health professions") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
ahp_statehospital_new <- statehospital_wte %>%
  filter(job_family == "Allied health professions") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
ahp_statehospital_prev_ndiff <- ahp_statehospital_prevquarter - ahp_statehospital_07
ahp_statehospital_new_ndiff <- ahp_statehospital_new - ahp_statehospital_07

# calculate the percentage difference with 2006
ahp_statehospital_prev_pdiff <- (ahp_statehospital_prev_ndiff/ahp_statehospital_07) *100
ahp_statehospital_new_pdiff <- (ahp_statehospital_new_ndiff/ahp_statehospital_07) *100

### For other State Hospital staff
# For 2007 figure
others_statehospital_07 <- statehospital_wte %>%
  filter(!job_family %in% c("Allied health professions", "Nursing & midwifery", "Medical & dental")) %>%
  summarise(sep07_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep07_sum)

# For previous quarter
others_statehospital_prevquarter <- statehospital_wte %>%
  filter(!job_family %in% c("Allied health professions", "Nursing & midwifery", "Medical & dental")) %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
others_statehospital_new <- statehospital_wte %>%
  filter(!job_family %in% c("Allied health professions", "Nursing & midwifery", "Medical & dental")) %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
others_statehospital_prev_ndiff <- others_statehospital_prevquarter - others_statehospital_07
others_statehospital_new_ndiff <- others_statehospital_new - others_statehospital_07

# calculate the percentage difference with 2006
others_statehospital_prev_pdiff <- (others_statehospital_prev_ndiff/others_statehospital_07) *100
others_statehospital_new_pdiff <- (others_statehospital_new_ndiff/others_statehospital_07) *100
```

**Previous: Sep-07 to `r prev_quarter`**  

* The overall workforce at State Hospital has `r ifelse ((statehospital_prev_pdiff<0),"decreased","increased")` by `r round_half_up(statehospital_prev_pdiff,1)`% or `r round_half_up(statehospital_prev_ndiff,1)` WTE, from `r round_half_up(statehospital_07,1)` WTE at September 2007 to `r round_half_up(statehospital_prevquarter,1)` WTE at `r prev_quarter`.  
* The Medical workforce at State Hospital has `r ifelse ((md_statehospital_prev_pdiff<0),"decreased","increased")` by  `r round_half_up(md_statehospital_prev_pdiff,1)`% or `r round_half_up(md_statehospital_prev_ndiff,1)` WTE, from `r round_half_up(md_statehospital_07,1)` WTE at September 2007 to `r round_half_up(md_statehospital_prevquarter,1)` WTE at `r prev_quarter`.  
*	The Nursing and Midwifery workforce at State Hospital has  `r ifelse ((nm_statehospital_prev_pdiff<0),"decreased","increased")` by  `r round_half_up(nm_statehospital_prev_pdiff,1)`% or `r round_half_up(nm_statehospital_prev_ndiff,1)` WTE, from `r round_half_up(nm_statehospital_07,1)` WTE at September 2007 to `r round_half_up(nm_statehospital_prevquarter,1)` WTE at `r prev_quarter`.  
*	The AHP workforce at State Hospital has `r ifelse ((ahp_statehospital_prev_pdiff<0),"decreased","increased")` by  `r round_half_up(ahp_statehospital_prev_pdiff,1)`% or `r round_half_up(ahp_statehospital_prev_ndiff,1)` WTE, from `r round_half_up(ahp_statehospital_07,1)` WTE at September 2007 to `r round_half_up(ahp_statehospital_prevquarter,1)` WTE at `r prev_quarter`. The number of AHPs, specifically occupational therapists, decreased suddenly in April 2012 at the time of the new State Hospital building opening. This decrease is likely to be the result of a recording/classification issue.  
*	Other staff at State Hospital have  `r ifelse ((others_statehospital_prev_pdiff<0),"decreased","increased")` by `r round_half_up(others_statehospital_prev_pdiff,1)`% or `r round_half_up(others_statehospital_prev_ndiff,1)` WTE, from `r round_half_up(others_statehospital_07,1)` WTE at September 2007 to `r round_half_up(others_statehospital_prevquarter,1)` WTE at `r prev_quarter`.  

**New: Sep-07 to `r new_quarter`**  

* The overall workforce at State Hospital has `r ifelse ((statehospital_new_pdiff<0),"decreased","increased")` by `r round_half_up(statehospital_new_pdiff,1)`% or `r round_half_up(statehospital_new_ndiff,1)` WTE, from `r round_half_up(statehospital_07,1)` WTE at September 2007 to `r round_half_up(statehospital_new,1)` WTE at `r new_quarter`.  
* The Medical workforce at State Hospital has `r ifelse ((md_statehospital_new_pdiff<0),"decreased","increased")` by  `r round_half_up(md_statehospital_new_pdiff,1)`% or `r round_half_up(md_statehospital_new_ndiff,1)` WTE, from `r round_half_up(md_statehospital_07,1)` WTE at September 2007 to `r round_half_up(md_statehospital_new,1)` WTE at `r new_quarter`.  
*	The Nursing and Midwifery workforce at State Hospital has  `r ifelse ((nm_statehospital_new_pdiff<0),"decreased","increased")` by  `r round_half_up(nm_statehospital_new_pdiff,1)`% or `r round_half_up(nm_statehospital_new_ndiff,1)` WTE, from `r round_half_up(nm_statehospital_07,1)` WTE at September 2007 to `r round_half_up(nm_statehospital_new,1)` WTE at `r new_quarter`.  
*	The AHP workforce at State Hospital has `r ifelse ((ahp_statehospital_new_pdiff<0),"decreased","increased")` by  `r round_half_up(ahp_statehospital_new_pdiff,1)`% or `r round_half_up(ahp_statehospital_new_ndiff,1)` WTE, from `r round_half_up(ahp_statehospital_07,1)` WTE at September 2007 to `r round_half_up(ahp_statehospital_new,1)` WTE at `r new_quarter`. The number of AHPs, specifically occupational therapists, decreased suddenly in April 2012 at the time of the new State Hospital building opening. This decrease is likely to be the result of a recording/classification issue.  
*	Other staff at State Hospital have  `r ifelse ((others_statehospital_new_pdiff<0),"decreased","increased")` by `r round_half_up(others_statehospital_new_pdiff,1)`% or `r round_half_up(others_statehospital_new_ndiff,1)` WTE, from `r round_half_up(others_statehospital_07,1)` WTE at September 2007 to `r round_half_up(others_statehospital_new,1)` WTE at `r new_quarter`.  