---
title: "FMQ A&E"
author: "Workforce and Pay Analytics"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```


```{r, include=TRUE, echo=FALSE}

### 0 - Setup

## Run setup scripts where needed data and libraries are loaded
## output destination folder also created there

source(here::here("code", "00_setup.R"))

```


```{r, include=TRUE, echo=FALSE}
### 1 - Read in rds files that were prepared in setup file

consultant_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_consultant_wte.rds")),
  )
```


```{r, include=TRUE, echo=FALSE}
# specify and format the briefing rounds
new_quarter <- as.Date(paste0(briefing_round, "-01"), format = "%Y-%B-%d")
new_quarter <-format(new_quarter, format = "%B %Y")


prev_quarter <- as.Date(paste0(prevquarter_briefing_round, "-01"), format = "%Y-%B-%d")
prev_quarter <-format(prev_quarter, format = "%B %Y")
```


```{r, include=TRUE, echo=FALSE}
# Add in 2006 figure for A&E consultants
emergency_cons_06 <- 75.755

# summarise data
# for previous quarter
emergency_cons_prev <- consultant_wte %>%
  filter(subjob_family == "Emergency medicine") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

#for current quarter
emergency_cons_new <- consultant_wte %>%
  filter(subjob_family == "Emergency medicine") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
ae_prev_ndiff <- emergency_cons_prev - emergency_cons_06
ae_new_ndiff <- emergency_cons_new - emergency_cons_06

# calculate the percentage difference with 2006
ae_prev_pdiff <- (ae_prev_ndiff/emergency_cons_06) *100
ae_new_pdiff <- (ae_new_ndiff/emergency_cons_06) *100
```


**Previous quarter (`r prev_quarter`):**

A&E Staffing - Under the SNP, a `r round_half_up(ae_prev_pdiff, 1)`% increase in A&E consultants (up `r round_half_up(ae_prev_ndiff, 1)` WTE - from `r round_half_up(emergency_cons_06, 1)` WTE in September 2006 to `r round_half_up(emergency_cons_prev, 1)` WTE in `r prev_quarter`).


**New quarter (`r new_quarter`):**

A&E Staffing - Under the SNP, a `r round_half_up(ae_new_pdiff, 1)`% increase in A&E consultants (up `r round_half_up(ae_new_ndiff, 1)` WTE - from `r round_half_up(emergency_cons_06, 1)` WTE in September 2006 to `r round_half_up(emergency_cons_new, 1)` WTE in `r new_quarter`).




