---
title: "FMQ NHS 24"
author: "Workforce and Pay Analytics"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: word_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```


```{r, include=TRUE, echo=FALSE}
### 0 - Setup

## Run setup scripts where needed data and libraries are loaded
## output destination folder also created there

source(here::here("code", "00_setup.R"))

```


```{r, include=TRUE, echo=FALSE}
### 1 - Read in rds files that were prepared in setup file

nhs24_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_nhs24_wte.rds")),
  )

hc_wte_data <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_hc_wte_data.rds")),
  )
```


```{r, include=TRUE, echo=FALSE}

# For 2007 figure
nhs24_wte_07 <- nhs24_wte %>%
  summarise(sep07_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep07_sum)

# For previous year
nhs24_wte_prevyear <- nhs24_wte %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# For previous quarter
nhs24_wte_prevquarter <- nhs24_wte %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
nhs24_wte_new <- nhs24_wte %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
nhs24wte_prev_ndiff <- nhs24_wte_prevquarter - nhs24_wte_07
nhs24wte_new_ndiff <- nhs24_wte_new - nhs24_wte_07

# calculate the percentage difference with 2006
nhs24wte_prev_pdiff <- (nhs24wte_prev_ndiff/nhs24_wte_07) *100
nhs24wte_new_pdiff <- (nhs24wte_new_ndiff/nhs24_wte_07) *100

# calculate the difference and percentage with previous year
nhs24wte_prevyear_ndiff <- nhs24_wte_new - nhs24_wte_prevyear
nhs24wte_prevyear_pdiff <- (nhs24wte_prevyear_ndiff/nhs24_wte_prevyear) *100

# calculate the difference and percentage with previous quarter
nhs24wte_prevqtr_ndiff <- nhs24_wte_new - nhs24_wte_prevquarter
nhs24wte_prevqtr_pdiff <- (nhs24wte_prevqtr_ndiff/nhs24_wte_prevquarter) *100


# For headcount figures of previous and current quarter
nhs24_hc_prevquarter <- hc_wte_data %>%
  filter(board == "NHS 24") %>%
  filter(census_date %in% prevquarter_briefing_round) %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)


nhs24_hc_new <- hc_wte_data %>%
  filter(board == "NHS 24") %>%
  filter(census_date %in% briefing_round) %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)
```


**Previous quarter (`r prev_quarter`):**

* As at `r prev_quarter`, NHS 24 has a total of `r format(round_half_up(nhs24_wte_prevquarter, 1), big.mark = ",")` WTE (`r format(round_half_up(nhs24_hc_prevquarter, 1), big.mark = ",")` headcount) staff in post. Total staff in post in NHS 24 have `r ifelse((nhs24wte_prev_pdiff < 0), "decreased", "increased")` by `r round_half_up(nhs24wte_prev_ndiff, 1)` WTE (`r round_half_up(nhs24wte_prev_pdiff, 1)`%) since September 2007.


**New quarter (`r new_quarter`):**

* As at `r new_quarter`, NHS 24 has a total of `r format(round_half_up(nhs24_wte_new, 1), big.mark = ",")` WTE (`r format(round_half_up(nhs24_hc_new, 1), big.mark = ",")` headcount) staff in post. Total staff in post in NHS 24 have `r ifelse((nhs24wte_new_pdiff < 0), "decreased", "increased")` by `r round_half_up(nhs24wte_new_ndiff, 1)` WTE (`r round_half_up(nhs24wte_new_pdiff, 1)`%) since September 2007, and have `r ifelse((nhs24wte_prevqtr_pdiff < 0), "decreased", "increased")` by `r round_half_up(nhs24wte_prevqtr_ndiff, 1)` WTE (`r round_half_up(nhs24wte_prevqtr_pdiff, 1)`%) over the last quarter.
