---
title: "Ministerial briefing"
author: "Workforce and Pay Analytics"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```

```{r, include=TRUE, echo=FALSE}
### 0 - Setup

## Run setup scripts where needed data and libraries are loaded
## output destination folder also created there

source(here::here("code", "00_setup.R"))

```


```{r, include=TRUE, echo=FALSE}
### 1 - Read in rds files that were prepared in setup file

detailed_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_detailed_wte.rds")),
  )

hc_wte_data <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_hc_wte_data.rds")),
  )

wte_pd <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_wte_pd.rds")),
  )

nm_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_nm_wte.rds")),
  )

consultantwte_td <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_consultantwte_td.rds")),
  )


### 1B. Check for the record highs
recordhigh_all <- hc_wte_data %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  slice(which.max(value)) %>%
  dplyr::pull(value)

recordhigh_nm <- hc_wte_data %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  slice(which.max(value)) %>%
  dplyr::pull(value)

recordhigh_ahp <- hc_wte_data %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Allied health professions") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  slice(which.max(value)) %>%
  dplyr::pull(value)

recordhigh_consultants <- consultantwte_td %>%
  slice(which.max(value)) %>%
  dplyr::pull(value)

```


```{r, include=TRUE, echo=FALSE}

### For ministerial briefing, we include headcount and wte for all staff, N&M, M&D  
# We also need some wte for specific specialties 

### For headcount and wte data for present quarter

allhc_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)

allwte_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)


### For all WTE data for other areas of interests
# 2006 wte data 
allwte_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# 2006 headcount
allhc_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)

# For reference to pre-pandemic, we use Dec 2019 data
allwte_dec19 <- hc_wte_data %>%
  filter(census_date == "2019-December") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# For previous year's total staff in post WTE
allwte_prevyear <- hc_wte_data %>%
  filter(census_date %in% prevyear_briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)


##### Comparisons of all staff data to different time points
# calculate the difference with 2006 and previous year
allwte_new_ndiff <- allwte_new - allwte_06
allhc_new_ndiff <- allhc_new - allhc_06

allwte_prevyear_ndiff <- allwte_new - allwte_prevyear
allwte_dec19_ndiff <- allwte_new - allwte_dec19
# calculate the percentage difference with 2006 and previous year
allwte_new_pdiff <- (allwte_new_ndiff/allwte_06) *100
allhc_new_pdiff <- (allhc_new_ndiff/allhc_06) *100

allwte_prevyear_pdiff <- (allwte_prevyear_ndiff/allwte_prevyear) *100
allwte_dec19_pdiff <- (allwte_dec19_ndiff/allwte_dec19) *100


##### For permanent wte
# For previous year
allpermanentwte_prevyear<- wte_pd %>%
  filter(contract == "Permanent") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# For previous quarter
allpermanentwte_prevquarter<- wte_pd %>%
  filter(contract == "Permanent") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

# For current quarter
allpermanentwte_new<- wte_pd %>%
  filter(contract == "Permanent") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with previous year/quarter
allpermanentwte_prevyear_ndiff <- allpermanentwte_new - allpermanentwte_prevyear
allpermanentwte_prevquarter_ndiff <- allpermanentwte_new - allpermanentwte_prevquarter
# calculate the percentage difference
allpermanentwte_prevyear_pdiff <- (allpermanentwte_prevyear_ndiff/allpermanentwte_prevyear) *100
allpermanentwte_prevquarter_pdiff <- (allpermanentwte_prevquarter_ndiff/allpermanentwte_prevquarter) *100


# For N&M analysis
#### For all N&M 2006 ####
nmwte_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

nmhc_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)

# For previous year's N&M WTE
nmwte_prevyear <- hc_wte_data %>%
  filter(census_date %in% prevyear_briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# For current quarter
nmwte_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

nmhc_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)

# calculate the difference with 2006
nmwte_new_ndiff <- nmwte_new - nmwte_06
nmhc_new_ndiff <- nmhc_new - nmhc_06
nm_prevyear_ndiff <- nmwte_new - nmwte_prevyear
# calculate the percentage difference with 2006
nmwte_new_pdiff <- (nmwte_new_ndiff/nmwte_06) *100
nmhc_new_pdiff <- (nmhc_new_ndiff/nmhc_06) *100
nm_prevyear_pdiff <- (nm_prevyear_ndiff/nmwte_prevyear) *100

#### For qualified N&M ####
# Since we don't have data for 2006 qualified N&M, we use 2007 data
qual_nmwte_06 <- 41026.2 #This figure is from legacy table/code value, deduced from previous publications. 

qual_nmwte_07 <- detailed_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(band_grouping == "Qualified") %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

# For current quarter for qual N&M
qual_nmwte_new <- detailed_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(band_grouping == "Qualified") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
qualnm_new_ndiff <- qual_nmwte_new - qual_nmwte_06
# calculate the percentage difference with 2007
qualnm_new_pdiff <- (qualnm_new_ndiff/qual_nmwte_06) *100


#### For qualified midwives ####
# Since we don't have data for 2006 N&M breakdown by band, we use 2007 data
#Figure from legacy table/code value can't not be verified

# Add in 2006 figure for qualified midwives
#qual_midwives_06 <- 2798.962 #I can't verify this figure. Took it from old Scotland table

qual_midwives_07 <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

# for previous year
qual_midwives_prevyear <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# for previous quarter
qual_midwives_prevquarter <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

#for current quarter
qual_midwives_new <- nm_wte %>%
 filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
qualmidwives_prev_ndiff <- qual_midwives_prevquarter - qual_midwives_07
qualmidwives_new_ndiff <- qual_midwives_new - qual_midwives_07

# calculate the percentage difference with 2007
qualmidwives_prev_pdiff <- (qualmidwives_prev_ndiff/qual_midwives_07) *100
qualmidwives_new_pdiff <- (qualmidwives_new_ndiff/qual_midwives_07) *100

# calculate the difference and percentage with previous year
qualmidwives_prevyear_ndiff <- qual_midwives_new - qual_midwives_prevyear
qualmidwives_prevyear_pdiff <- (qualmidwives_prevyear_ndiff/qual_midwives_prevyear) *100

# calculate the difference and percentage with previous quarter
qualmidwives_prevqtr_ndiff <- qual_midwives_new - qual_midwives_prevquarter
qualmidwives_prevqtr_pdiff <- (qualmidwives_prevqtr_ndiff/qual_midwives_prevquarter) *100


#### For mental health nurses ####
# for previous year
mh_nurses_prevyear <- nm_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(subjob_family == "Mental health") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# for previous quarter
mh_nurses_prevquarter <- nm_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(subjob_family == "Mental health") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

#for current quarter
mh_nurses_new <- nm_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(subjob_family == "Mental health") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)


# calculate the difference and percentage with previous year
mhnurses_prevyear_ndiff <- mh_nurses_new - mh_nurses_prevyear
mhnurses_prevyear_pdiff <- (mhnurses_prevyear_ndiff/mh_nurses_prevyear) *100

# calculate the difference and percentage with previous quarter
mhnurses_prevqtr_ndiff <- mh_nurses_new - mh_nurses_prevquarter
mhnurses_prevqtr_pdiff <- (mhnurses_prevqtr_ndiff/mh_nurses_prevquarter) *100


#### For AHPs (excluding paramedics)####

# For 2006 data - can use the headcount/wte data because paramedics were 
# not included in AHPs in 2006
ahpwte_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Allied health professions") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# We have to use the detailed data so that we can remove 
# paramedics from them

# For current quarter
ahpwte_prevyear <- detailed_wte %>%
  filter(job_family == "Allied health professions") %>%
  filter(!subjob_family == "Paramedics") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# For current quarter
ahpwte_new <- detailed_wte %>%
  filter(job_family == "Allied health professions") %>%
  filter(!subjob_family == "Paramedics") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006 and previous year
ahp_new_ndiff <- ahpwte_new - ahpwte_06
ahp_prevyear_ndiff <- ahpwte_new - ahpwte_prevyear
# calculate the percentage difference with 2006
ahp_new_pdiff <- (ahp_new_ndiff/ahpwte_06) *100
ahp_prevyear_pdiff <- (ahp_prevyear_ndiff/ahpwte_prevyear) *100


#### For paramedics ####
paramedics_07 <- detailed_wte %>%
  filter(subjob_family == "Paramedics") %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

paramedics_prevyear <- detailed_wte %>%
  filter(job_family == "Allied health professions") %>%
  filter(subjob_family == "Paramedics") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)


paramedics_new <- detailed_wte %>%
  filter(job_family == "Allied health professions") %>%
  filter(subjob_family == "Paramedics") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
paramedics_new_ndiff <- paramedics_new - paramedics_07
paramedics_prevyear_ndiff <- paramedics_new - paramedics_prevyear
# calculate the percentage difference with 2006
paramedics_new_pdiff <- (paramedics_new_ndiff/paramedics_07) *100
paramedics_prevyear_pdiff <- (paramedics_prevyear_ndiff/paramedics_prevyear) *100



#### For all M&D workforce ####
# For previous year's M&D WTE
mdwte_prevyear <- hc_wte_data %>%
  filter(census_date %in% prevyear_briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Medical & dental") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# For current quarter
mdwte_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Medical & dental") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)


# calculate the difference with 2006
md_prevyear_ndiff <- mdwte_new - mdwte_prevyear
# calculate the percentage difference with 2006
md_prevyear_pdiff <- (md_prevyear_ndiff/mdwte_prevyear) *100


####### For relevant data of specific consultant specialties ####
# We use data that can breakdown by bands or sub-job families

# Consultant wte for 2006 are hard coded because data table not available on 
# new dashboards. Hard coded data are taken from ISD tables.

#### For all consultants ####
consultants_wte_06 <- 3636.545

consultants_wte_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

consultants_wte_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
consultants_new_ndiff <- consultants_wte_new - consultants_wte_06
consultants_prevyear_ndiff <- consultants_wte_new - consultants_wte_prevyear
# calculate the percentage difference with 2006
consultants_new_pdiff <- (consultants_new_ndiff/consultants_wte_06) *100
consultants_prevyear_pdiff <- (consultants_prevyear_ndiff/consultants_wte_prevyear) *100


#### For Radiology consultants ####
radiology_cons_06 <- 223.525

radiology_cons_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Clinical radiology") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

radiology_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Clinical radiology") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006 and previous year
radiology_new_ndiff <- radiology_cons_new - radiology_cons_06
radiology_prevyear_ndiff <- radiology_cons_new - radiology_cons_prevyear
# calculate the percentage difference with 2006 and previous year
radiology_new_pdiff <- (radiology_new_ndiff/radiology_cons_06) *100
radiology_prevyear_pdiff <- (radiology_prevyear_ndiff/radiology_cons_prevyear) *100


#### For Paediatrics consultants ####
paediatrics_cons_06 <- 162.972

paediatrics_cons_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Paediatrics") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

paediatrics_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Paediatrics") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
paediatrics_new_ndiff <- paediatrics_cons_new - paediatrics_cons_06
paediatrics_prevyear_ndiff <- paediatrics_cons_new - paediatrics_cons_prevyear
# calculate the percentage difference with 2006
paediatrics_new_pdiff <- (paediatrics_new_ndiff/paediatrics_cons_06) *100
paediatrics_prevyear_pdiff <- (paediatrics_prevyear_ndiff/paediatrics_cons_prevyear) *100


#### For A&E consultants ####
emergency_cons_06 <- 75.76

emergency_cons_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Emergency medicine") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

emergency_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Emergency medicine") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
emergency_new_ndiff <- emergency_cons_new - emergency_cons_06
emergency_prevyear_ndiff <- emergency_cons_new - emergency_cons_prevyear
# calculate the percentage difference with 2006
emergency_new_pdiff <- (emergency_new_ndiff/emergency_cons_06) *100
emergency_prevyear_pdiff <- (emergency_prevyear_ndiff/emergency_cons_prevyear) *100

#### For all Psychiatry consultants ####
psychiatry_cons_06 <- 444.668

psychiatry_cons_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family %in% c("General psychiatry", "Child & adolescent psychiatry", "Forensic psychiatry", "Old age psychiatry", "Psychiatry of learning disability", "Psychotherapy")) %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

psychiatry_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family %in% c("General psychiatry", "Child & adolescent psychiatry", "Forensic psychiatry", "Old age psychiatry", "Psychiatry of learning disability", "Psychotherapy")) %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
psychiatry_new_ndiff <- psychiatry_cons_new - psychiatry_cons_06
psychiatry_prevyear_ndiff <- psychiatry_cons_new - psychiatry_cons_prevyear
# calculate the percentage difference with 2006
psychiatry_new_pdiff <- (psychiatry_new_ndiff/psychiatry_cons_06) *100
psychiatry_prevyear_pdiff <- (psychiatry_prevyear_ndiff/psychiatry_cons_prevyear) *100


#### For Oncology consultants ####
oncology_cons_06 <- 69.542

oncology_cons_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family %in% c("Clinical oncology", "Medical oncology")) %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

oncology_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family %in% c("Clinical oncology", "Medical oncology")) %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
oncology_new_ndiff <- oncology_cons_new - oncology_cons_06
oncology_prevyear_ndiff <- oncology_cons_new - oncology_cons_prevyear
# calculate the percentage difference with 2006
oncology_new_pdiff <- (oncology_new_ndiff/oncology_cons_06) *100
oncology_prevyear_pdiff <- (oncology_prevyear_ndiff/oncology_cons_prevyear) *100



#### For medical & dental workforce
# Current quarter
mdwte_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Medical & dental") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# For previous quarter M&D workforce
mdwte_prevyear <- hc_wte_data %>%
  filter(census_date %in% prevyear_briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Medical & dental") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)


```

__NHS Workforce - Headlines__

As of `r new_quarter`,

__Overall picture__

* About `r format(round(allhc_new, -2), big.mark = ",")` staff are employed by NHS Scotland (`r format(allhc_new, big.mark = ",")` headcount). This is equivalent to about `r format(round(allwte_new, -2), big.mark = ",")` WTE, which has __`r ifelse((allwte_prevyear_pdiff < 0), "decreased", "increased")` by `r round_half_up(allwte_prevyear_pdiff, 1)`%__ over the last year (by `r format(round_half_up(allwte_prevyear_ndiff, 1), big.mark =",")` to `r format(round_half_up(allwte_new, 1), big.mark = ",")` WTE).

__Trend__

* Since September 2006, the number of NHS Scotland staff has `r ifelse((allwte_new_pdiff < 0), "decreased", "increased")` by `r format(allhc_new_ndiff, big.mark = ",")` (`r format(round_half_up(allwte_new_ndiff, 1), big.mark = ",")` WTE), `r ifelse((allhc_new_pdiff < 0), "decreasing", "increasing")` the headcount by `r round_half_up(allhc_new_pdiff, 1)`% (`r round_half_up(allwte_new_pdiff, 1)`%) from a headcount of `r format(allhc_06, big.mark = ",")` (`r format(round_half_up(allwte_06, 1), big.mark = ",")` WTE) in 2006.

__Nursing and Midwifery__

* There are `r ifelse ((round(nmhc_new, -3) > nmhc_new),"almost,","over")` `r format(round(nmhc_new, -3), big.mark = ",")` (`r format(nmhc_new, big.mark = ",")` headcount) __Nursing and Midwifery staff__ in NHS Scotland workforce, an increase of `r format(nmhc_new_ndiff, big.mark = ",")` headcount (`r round_half_up(nmhc_new_pdiff, 1)`%) since September 2006. This is equivalent to `r ifelse ((round(nmhc_new, -3) > nmhc_new),"almost,","over")` `r format(round(nmwte_new, -3), big.mark = ",")` (`r format(round_half_up(nmwte_new, 1), big.mark = ",")` WTE) staff, `r ifelse((nm_prevyear_pdiff < 0), "a decrease", "an increase")` of `r format(round_half_up(nm_prevyear_pdiff, 1), big.mark = ",")`% (`r ifelse ((nm_prevyear_ndiff > 0),"+","")``r round_half_up(nm_prevyear_ndiff, 1)` WTE) since last year, and `r ifelse((nmwte_new_pdiff < 0), "a decrease", "an increase")` of `r format(round_half_up(nmwte_new_ndiff, 1), big.mark = ",")` WTE (`r round_half_up(nmwte_new_pdiff, 1)`%) since September 2006.

*	There are `r ifelse ((round(qual_nmwte_new, -3) > qual_nmwte_new),"almost","over")` `r format(round(qual_nmwte_new, -3), big.mark = ",")` (`r format(round_half_up(qual_nmwte_new, 1), big.mark =",")` WTE) __qualified nurses and midwives__ working in NHS Scotland, a `r round_half_up(qualnm_new_pdiff, 1)`%  increase of `r ifelse ((round(qualnm_new_ndiff, -3) > qualnm_new_ndiff),"almost","over")` `r format(round(qualnm_new_ndiff, -3), big.mark = ",")` (`r format(round_half_up(qualnm_new_ndiff, 1), big.mark =",")` WTE) since September 2006.  

* There are `r ifelse ((round(qual_midwives_new, -2) > qual_midwives_new),"almost,","over")` `r format(round(qual_midwives_new, -2), big.mark = ",")` (`r format(round_half_up(qual_midwives_new, 1), big.mark =",")` WTE) __qualified midwives__ working in NHS Scotland, a __`r round_half_up(qualmidwives_prevyear_pdiff, 1)`% `r ifelse((qualmidwives_prevyear_pdiff < 0), "decrease", "increase")`__ (`r ifelse ((qualmidwives_prevyear_ndiff > 0),"+","")``r round_half_up(qualmidwives_prevyear_ndiff, 1)` WTE) over the last year, but a __`r round_half_up(qualmidwives_prevqtr_pdiff, 1)`% (`r round_half_up(qualmidwives_prevqtr_ndiff, 1)` WTE) `r ifelse((qualmidwives_prevqtr_pdiff < 0), "decrease", "increase")`__ over the last quarter. 


__Medical and Dental__

*	There are `r format(round_half_up(consultants_wte_new, 1), big.mark = ",")` WTE __medical & dental consultants__ (including directors) in post, `r ifelse ((consultants_wte_new > consultants_wte_prevyear),"an increase","a decrease")` of `r round_half_up(consultants_prevyear_pdiff, 1)`% (`r ifelse ((consultants_prevyear_ndiff > 0),"+","")``r round_half_up(consultants_prevyear_ndiff, 1)` WTE) since last year, and `r ifelse ((consultants_wte_new > consultants_wte_06),"an increase","a decrease")` of `r round_half_up(consultants_new_pdiff, 1)`% (`r ifelse ((consultants_new_ndiff > 0),"+","")``r format(round_half_up(consultants_new_ndiff, 1), big.mark =",")` WTE) since September 2006.

* There are `r round_half_up(radiology_cons_new, 1)` WTE __clinical radiology consultants__ working in NHS Scotland, `r ifelse (radiology_cons_new > radiology_cons_06,"an increase","a decrease")` of `r round_half_up(radiology_new_pdiff, 1)`% (`r ifelse ((radiology_new_ndiff > 0),"+","")``r round_half_up(radiology_new_ndiff, 1)` WTE) since September 2006, and `r ifelse (radiology_cons_new > radiology_cons_prevyear,"an increase,","a decrease")` of `r round_half_up(radiology_prevyear_pdiff, 1)`% (`r ifelse ((radiology_prevyear_ndiff > 0),"+","")``r round_half_up(radiology_prevyear_ndiff, 1)` WTE) over the last year.

* There are `r round_half_up(paediatrics_cons_new, 1)` WTE __paediatric consultants__ working in NHS Scotland, `r ifelse (paediatrics_cons_new > paediatrics_cons_06,"an increase","a decrease")` of `r round_half_up(paediatrics_new_pdiff, 1)`% (`r ifelse ((paediatrics_new_ndiff > 0),"+","")``r round_half_up(paediatrics_new_ndiff, 1)` WTE) since September 2006, and `r ifelse (paediatrics_cons_new > paediatrics_cons_prevyear,"an increase","a decrease")` of `r round_half_up(paediatrics_prevyear_pdiff, 1)`% (`r ifelse ((paediatrics_prevyear_ndiff > 0),"+","")``r round_half_up(paediatrics_prevyear_ndiff, 1)` WTE) over the last year.

__Allied Health Professionals (AHPs) (excl Paramedics)__

*	There are `r format(round_half_up(ahpwte_new, 1), big.mark =",")` WTE AHPs (excluding paramedics) working in NHS Scotland. This represents `r ifelse (ahpwte_new > ahpwte_prevyear,"an annual increase","an annual decrease")` of `r round_half_up(ahp_prevyear_pdiff, 1)`% (`r format(round_half_up(ahp_prevyear_ndiff,1), big.mark =",")` WTE), and `r ifelse (ahp_new_pdiff > 0,"an overall increase,","an overall decrease")` of `r round_half_up(ahp_new_pdiff, 1)`% since September 2006.


Optional lines:

* There are `r round_half_up(emergency_cons_new, 1)` WTE __Emergency Medicine consultants__ working in NHS Scotland, `r ifelse (emergency_cons_new > emergency_cons_06,"an increase","a decrease")` of `r round_half_up(emergency_new_pdiff, 1)`% (`r ifelse ((emergency_new_ndiff > 0),"+","")``r round_half_up(emergency_new_ndiff, 1)` WTE) since September 2006, and `r ifelse (emergency_cons_new > emergency_cons_prevyear,"an increase","a decrease")` of `r round_half_up(emergency_prevyear_pdiff, 1)`% (`r ifelse ((emergency_prevyear_ndiff > 0),"+","")``r round_half_up(emergency_prevyear_ndiff, 1)` WTE) over the last year.

* There are `r round_half_up(oncology_cons_new, 1)` WTE __consultant Oncologists__ working in NHS Scotland, `r ifelse (oncology_cons_new > oncology_cons_06,"an increase","a decrease")` of `r round_half_up(oncology_new_pdiff, 1)`% (`r ifelse ((oncology_new_ndiff > 0),"+","")``r round_half_up(oncology_new_ndiff, 1)` WTE) since September 2006, and `r ifelse (oncology_cons_new > oncology_cons_prevyear,"an increase","a decrease")` of `r round_half_up(oncology_prevyear_pdiff, 1)`% (`r ifelse ((oncology_prevyear_ndiff > 0),"+","")``r round_half_up(oncology_prevyear_ndiff, 1)` WTE) over the last year.

* There are `r round_half_up(psychiatry_cons_new, 1)` WTE __Psychiatric specialties consultants__ working in NHS Scotland, `r ifelse (psychiatry_cons_new > psychiatry_cons_06,"an increase","a decrease")` of `r round_half_up(psychiatry_new_pdiff, 1)`% (`r ifelse ((psychiatry_new_ndiff > 0),"+","")``r round_half_up(psychiatry_new_ndiff, 1)` WTE) since September 2006, and `r ifelse (psychiatry_cons_new > psychiatry_cons_prevyear,"an increase","a decrease")` of `r round_half_up(psychiatry_prevyear_pdiff, 1)`% (`r ifelse ((psychiatry_prevyear_ndiff > 0),"+","")``r round_half_up(psychiatry_prevyear_ndiff, 1)` WTE) over the last year.

* There are `r format(round_half_up(mh_nurses_new, 1), big.mark =",")` WTE __mental health nurses__ working in NHS Scotland, a `r round_half_up(mhnurses_prevyear_pdiff, 1)`% (`r ifelse ((mhnurses_prevyear_ndiff > 0),"+","")``r round_half_up(mhnurses_prevyear_ndiff, 1)` WTE) `r ifelse (mh_nurses_new > mh_nurses_prevyear,"increase","decrease")` over the last year, but a `r round_half_up(mhnurses_prevqtr_pdiff, 1)`% (`r ifelse ((mhnurses_prevqtr_ndiff > 0),"+","")``r round_half_up(mhnurses_prevqtr_ndiff, 1)` WTE) `r ifelse (mh_nurses_new > mh_nurses_prevquarter,"increase","decrease")` over the last quarter.

* There are `r format(round_half_up(paramedics_new, 1), big.mark =",")` WTE __paramedics__ working in NHS Scotland. This represents an annual `r ifelse (paramedics_new > paramedics_prevyear,"increase","decrease")` of `r round_half_up(paramedics_prevyear_pdiff, 1)`%  (`r ifelse ((paramedics_prevyear_ndiff > 0),"+","")``r round_half_up(paramedics_prevyear_ndiff, 1)` WTE), and an overall `r ifelse (paramedics_new > paramedics_07,"increase","decrease")` of `r round_half_up(paramedics_new_pdiff, 1)`% since September 2007.

* The employed medical and dental workforce has seen a `r round_half_up(md_prevyear_pdiff, 1)` % growth year-on-year (`r prev_year` - `r new_quarter`).
