---
title: "Workforce Core brief"
author: "Workforce and Pay Analytics"
date: "`r format(Sys.Date(), '%d %B %Y')`"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning=FALSE)
```

```{r, include=TRUE, echo=FALSE}
### 0 - Setup

## Run setup scripts where needed data and libraries are loaded
## output destination folder also created there

source(here::here("code", "00_setup.R"))

```


```{r, include=TRUE, echo=FALSE}
### 1 - Read in rds files that were prepared in setup file

detailed_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_detailed_wte.rds")),
  )

hc_wte_data <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_hc_wte_data.rds")),
  )

wte_pd <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_wte_pd.rds")),
  )

nm_wte <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_nm_wte.rds")),
  )

consultantwte_td <- read_rds(
  here("data", folder_loc, paste0(briefing_round, "_consultantwte_td.rds")),
  )


### 1B. Check for the record highs
recordhigh_all <- hc_wte_data %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  slice(which.max(value)) %>%
  dplyr::pull(value)

recordhigh_nm <- hc_wte_data %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  slice(which.max(value)) %>%
  dplyr::pull(value)

recordhigh_ahp <- hc_wte_data %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Allied health professions") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  slice(which.max(value)) %>%
  dplyr::pull(value)

recordhigh_consultants <- consultantwte_td %>%
  slice(which.max(value)) %>%
  dplyr::pull(value)
```


```{r, include=TRUE, echo=FALSE}

### For this brief, we include headcount and wte for all staff, N&M, M&D  
# We also need some wte for specific specialties 

### For headcount and wte data for present quarter

allhc_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)

allwte_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)


### For all WTE data for other areas of interests
# 2006 wte data 
allwte_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# 2006 headcount
allhc_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)

# For reference to pre-pandemic, we use Dec 2019 data
allwte_dec19 <- hc_wte_data %>%
  filter(census_date == "2019-December") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# For previous year's total staff in post WTE
allwte_prevyear <- hc_wte_data %>%
  filter(census_date %in% prevyear_briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "All staff") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)


##### Comparisons of all staff data to different time points
# calculate the difference with 2006 and previous year
allwte_new_ndiff <- allwte_new - allwte_06
allhc_new_ndiff <- allhc_new - allhc_06

allwte_prevyear_ndiff <- allwte_new - allwte_prevyear
allwte_dec19_ndiff <- allwte_new - allwte_dec19
# calculate the percentage difference with 2006 and previous year
allwte_new_pdiff <- (allwte_new_ndiff/allwte_06) *100
allhc_new_pdiff <- (allhc_new_ndiff/allhc_06) *100

allwte_prevyear_pdiff <- (allwte_prevyear_ndiff/allwte_prevyear) *100
allwte_dec19_pdiff <- (allwte_dec19_ndiff/allwte_dec19) *100


##### For permanent wte
# For previous year
allpermanentwte_prevyear<- wte_pd %>%
  filter(contract == "Permanent") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# For current quarter
allpermanentwte_new<- wte_pd %>%
  filter(contract == "Permanent") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with previous year
allpermanentwte_prevyear_ndiff <- allpermanentwte_new - allpermanentwte_prevyear
# calculate the percentage difference
allpermanentwte_prevyear_pdiff <- (allpermanentwte_prevyear_ndiff/allpermanentwte_prevyear) *100


##### For N&M analysis
# For all N&M 2006
nmwte_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

nmhc_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)

# For previous year's N&M WTE
nmwte_prevyear <- hc_wte_data %>%
  filter(census_date %in% prevyear_briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# For current quarter
nmwte_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

nmhc_new <- hc_wte_data %>%
  filter(census_date %in% briefing_round) %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(headcount_wte == "Headcount") %>%
  dplyr::pull(value)

# calculate the difference with 2006
nmwte_new_ndiff <- nmwte_new - nmwte_06
nmhc_new_ndiff <- nmhc_new - nmhc_06
nm_prevyear_ndiff <- nmwte_new - nmwte_prevyear
# calculate the percentage difference with 2006
nmwte_new_pdiff <- (nmwte_new_ndiff/nmwte_06) *100
nmhc_new_pdiff <- (nmhc_new_ndiff/nmhc_06) *100
nm_prevyear_pdiff <- (nm_prevyear_ndiff/nmwte_prevyear) *100

##### For qualified N&M
# Since we don't have data for 2006 qualified N&M, we use 2007 data
#qual_nmwte_06 <- 41026.2 The figure is from legacy table/code value and could not be verified

qual_nmwte_07 <- detailed_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(band_grouping == "Qualified") %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

# For current quarter for qual N&M
qual_nmwte_new <- detailed_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(band_grouping == "Qualified") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
qualnm_new_ndiff <- qual_nmwte_new - qual_nmwte_07
# calculate the percentage difference with 2007
qualnm_new_pdiff <- (qualnm_new_ndiff/qual_nmwte_07) *100


##### For qualified midwives
# Since we don't have data for 2006 N&M breakdown by band, we use 2007 data
#Figure from legacy table/code value can't not be verified

# Add in 2006 figure for qualified midwives
#qual_midwives_06 <- 2798.962 #I can't verify this figure. Took it from old Scotland table

qual_midwives_07 <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

# for previous year
qual_midwives_prevyear <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# for previous quarter
qual_midwives_prevquarter <- nm_wte %>%
  filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(prevquarter_sum = sum(prevquarter_wte)) %>%
  dplyr::pull(prevquarter_sum)

#for current quarter
qual_midwives_new <- nm_wte %>%
 filter(subjob_family %in% midwifery_specialties) %>%
  filter(band_grouping == "Qualified") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
qualmidwives_prev_ndiff <- qual_midwives_prevquarter - qual_midwives_07
qualmidwives_new_ndiff <- qual_midwives_new - qual_midwives_07

# calculate the percentage difference with 2007
qualmidwives_prev_pdiff <- (qualmidwives_prev_ndiff/qual_midwives_07) *100
qualmidwives_new_pdiff <- (qualmidwives_new_ndiff/qual_midwives_07) *100

# calculate the difference and percentage with previous year
qualmidwives_prevyear_ndiff <- qual_midwives_new - qual_midwives_prevyear
qualmidwives_prevyear_pdiff <- (qualmidwives_prevyear_ndiff/qual_midwives_prevyear) *100

# calculate the difference and percentage with previous quarter
qualmidwives_prevqtr_ndiff <- qual_midwives_new - qual_midwives_prevquarter
qualmidwives_prevqtr_pdiff <- (qualmidwives_prevqtr_ndiff/qual_midwives_prevquarter) *100


### For mental health nurses
mh_nmwte_07 <- detailed_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(subjob_family == "Mental health") %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

# For current quarter for qual N&M
mh_nmwte_new <- detailed_wte %>%
  filter(job_family == "Nursing & midwifery") %>%
  filter(subjob_family == "Mental health") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
mhnm_new_ndiff <- mh_nmwte_new - mh_nmwte_07
# calculate the percentage difference with 2007
mhnm_new_pdiff <- (mhnm_new_ndiff/mh_nmwte_07) *100



##### For AHPs (excluding paramedics)

# For 2006 data - can use the headcount/wte data because paramedics were 
# not included in AHPs in 2006
ahpwte_06 <- hc_wte_data %>%
  filter(census_date == "2006-September") %>%
  filter(board == "Scotland") %>%
  filter(job_family == "Allied health professions") %>%
  filter(headcount_wte == "Whole Time Equivalent") %>%
  dplyr::pull(value)

# We have to use the detailed data so that we can remove 
# paramedics from them

# For current quarter
ahpwte_prevyear <- detailed_wte %>%
  filter(job_family == "Allied health professions") %>%
  filter(!subjob_family == "Paramedics") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

# For current quarter
ahpwte_new <- detailed_wte %>%
  filter(job_family == "Allied health professions") %>%
  filter(!subjob_family == "Paramedics") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006 and previous year
ahp_new_ndiff <- ahpwte_new - ahpwte_06
ahp_prevyear_ndiff <- ahpwte_new - ahpwte_prevyear
# calculate the percentage difference with 2006
ahp_new_pdiff <- (ahp_new_ndiff/ahpwte_06) *100
ahp_prevyear_pdiff <- (ahp_prevyear_ndiff/ahpwte_prevyear) *100


##### For Paramedics
# Note that these were initially categorized under Ambulance Support Services 
# from 2007 until 2013 when they were grouped under AHPs

# For 2007
paramedics_07 <- detailed_wte %>%
  filter(subjob_family == "Paramedics") %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

# For current quarter
paramedics_new <- detailed_wte %>%
  filter(subjob_family == "Paramedics") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006 and previous year
paramedics_new_ndiff <- paramedics_new - paramedics_07

# calculate the percentage difference with 2006
paramedics_new_pdiff <- (paramedics_new_ndiff/paramedics_07) *100




####### For relevant data for specific specialties
# We use data that can breakdown by bands or sub-job families

# Consultant wte for 2006 are hard coded because data table not available on 
# new dashboards. Hard coded data are taken from ISD tables.

##### Now, For Medical consultant only
# For 2006 data
medical_wte_06 <- 9599.975

# For current quarter  
medical_wte_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(!subjob_family %in% dental_specialties) %>%
  summarise(new_sum = sum(new_wte)) %>% 
  dplyr::pull(new_sum)

# calculate the difference with 2006
medical_new_ndiff <- medical_wte_new - medical_wte_06

# calculate the percentage difference with 2006
medical_new_pdiff <- (medical_new_ndiff/medical_wte_06) *100


##### For doctors in training
# For 2006 data
doctorsintraining_wte_06 <- 5336.13

# For current quarter  
doctorsintraining_wte_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Doctor in training") %>%
  summarise(new_sum = sum(new_wte)) %>% 
  dplyr::pull(new_sum)

# calculate the difference with 2006
doctorsintraining_new_ndiff <- doctorsintraining_wte_new - doctorsintraining_wte_06

# calculate the percentage difference with 2006
doctorsintraining_new_pdiff <- (doctorsintraining_new_ndiff/doctorsintraining_wte_06) *100



##### For all consultants
consultants_wte_06 <- 3636.545

consultants_wte_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

consultants_wte_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
consultants_new_ndiff <- consultants_wte_new - consultants_wte_06
consultants_prevyear_ndiff <- consultants_wte_new - consultants_wte_prevyear
# calculate the percentage difference with 2006
consultants_new_pdiff <- (consultants_new_ndiff/consultants_wte_06) *100
consultants_prevyear_pdiff <- (consultants_prevyear_ndiff/consultants_wte_prevyear) *100


##### For A&E consultants
emergency_cons_06 <- 75.755

emergency_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Emergency medicine") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
emergency_new_ndiff <- emergency_cons_new - emergency_cons_06
# calculate the percentage difference with 2006
emergency_new_pdiff <- (emergency_new_ndiff/emergency_cons_06) *100


##### For Geriatric consultants
geriatric_cons_06 <- 118.905

geriatric_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Geriatric medicine") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
geriatric_new_ndiff <- geriatric_cons_new - geriatric_cons_06
# calculate the percentage difference with 2006
geriatric_new_pdiff <- (geriatric_new_ndiff/geriatric_cons_06) *100


##### For Respiratory medicine consultants
respiratory_cons_06 <- 68.47

respiratory_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Respiratory medicine") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
respiratory_new_ndiff <- respiratory_cons_new - respiratory_cons_06
# calculate the percentage difference with 2006
respiratory_new_pdiff <- (respiratory_new_ndiff/respiratory_cons_06) *100


##### For Infectious diseases consultants
infectiousdiseases_cons_06 <- 18.454

infectiousdiseases_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Infectious diseases") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
infectiousdiseases_new_ndiff <- infectiousdiseases_cons_new - infectiousdiseases_cons_06
# calculate the percentage difference with 2006
infectiousdiseases_new_pdiff <- (infectiousdiseases_new_ndiff/infectiousdiseases_cons_06) *100


##### For Oncology consultants
oncology_cons_06 <- 69.542

oncology_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family %in% c("Clinical oncology", "Medical oncology")) %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
oncology_new_ndiff <- oncology_cons_new - oncology_cons_06
# calculate the percentage difference with 2006
oncology_new_pdiff <- (oncology_new_ndiff/oncology_cons_06) *100


##### For Radiology consultants
radiology_cons_06 <- 223.525

radiology_cons_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Clinical radiology") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

radiology_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Clinical radiology") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006 and previous year
radiology_new_ndiff <- radiology_cons_new - radiology_cons_06
radiology_prevyear_ndiff <- radiology_cons_new - radiology_cons_prevyear
# calculate the percentage difference with 2006 and previous year
radiology_new_pdiff <- (radiology_new_ndiff/radiology_cons_06) *100
radiology_prevyear_pdiff <- (radiology_prevyear_ndiff/radiology_cons_prevyear) *100



##### For Paediatrics consultants
paediatrics_cons_06 <- 162.972

paediatrics_cons_prevyear <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Paediatrics") %>%
  summarise(prevyear_sum = sum(prevyear_wte)) %>%
  dplyr::pull(prevyear_sum)

paediatrics_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Paediatrics") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
paediatrics_new_ndiff <- paediatrics_cons_new - paediatrics_cons_06
paediatrics_prevyear_ndiff <- paediatrics_cons_new - paediatrics_cons_prevyear
# calculate the percentage difference with 2006
paediatrics_new_pdiff <- (paediatrics_new_ndiff/paediatrics_cons_06) *100
paediatrics_prevyear_pdiff <- (paediatrics_prevyear_ndiff/paediatrics_cons_prevyear) *100



##### For radiography staff
# For 2007
radiography_07 <- detailed_wte %>%
  filter(job_family == "Allied health professions") %>%
  filter(subjob_family %in% c("Radiography - diagnostic", "Radiography - theraputic")) %>%
  summarise(sep2007_sum = sum(sep2007_wte)) %>%
  dplyr::pull(sep2007_sum)

radiography_new <- detailed_wte %>%
  filter(job_family == "Allied health professions") %>%
  filter(subjob_family %in% c("Radiography - diagnostic", "Radiography - theraputic")) %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2007
radiography_new_ndiff <- radiography_new - radiography_07
# calculate the percentage difference with 2007
radiography_new_pdiff <- (radiography_new_ndiff/radiography_07) *100


##### For all Psychiatry consultants
psychiatry_cons_06 <- 444.668

psychiatry_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family %in% c("General psychiatry", "Child & adolescent psychiatry", "Forensic psychiatry", "Old age psychiatry", "Psychiatry of learning disability", "Psychotherapy")) %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
psychiatry_new_ndiff <- psychiatry_cons_new - psychiatry_cons_06
# calculate the percentage difference with 2006
psychiatry_new_pdiff <- (psychiatry_new_ndiff/psychiatry_cons_06) *100


# For General Psychiatry consultants
genpsychiatry_cons_06 <- 244.713

genpsychiatry_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "General psychiatry") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
genpsychiatry_new_ndiff <- genpsychiatry_cons_new - genpsychiatry_cons_06
# calculate the percentage difference with 2006
genpsychiatry_new_pdiff <- (genpsychiatry_new_ndiff/genpsychiatry_cons_06) *100


### For Child & Adolescent Psychiatry consultants
childpsychiatry_cons_06 <- 55.713

childpsychiatry_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Child & adolescent psychiatry") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
childpsychiatry_new_ndiff <- childpsychiatry_cons_new - childpsychiatry_cons_06
# calculate the percentage difference with 2006
childpsychiatry_new_pdiff <- (childpsychiatry_new_ndiff/childpsychiatry_cons_06) *100


# For Forensic Psychiatry consultants
forensicpsychiatry_cons_06 <- 35.828

forensicpsychiatry_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Forensic psychiatry") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
forensicpsychiatry_new_ndiff <- forensicpsychiatry_cons_new - forensicpsychiatry_cons_06
# calculate the percentage difference with 2006
forensicpsychiatry_new_pdiff <- (forensicpsychiatry_new_ndiff/forensicpsychiatry_cons_06) *100


# For Old age Psychiatry consultants
oldagepsychiatry_cons_06 <- 68.68

oldagepsychiatry_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Old age psychiatry") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
oldagepsychiatry_new_ndiff <- oldagepsychiatry_cons_new - oldagepsychiatry_cons_06
# calculate the percentage difference with 2006
oldagepsychiatry_new_pdiff <- (oldagepsychiatry_new_ndiff/oldagepsychiatry_cons_06) *100


# For Psychiatry of learning disability consultants
learningdisability_psychiatry_cons_06 <- 24.844

learningdisability_psychiatry_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Psychiatry of learning disability") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
learningdisability_psychiatry_new_ndiff <- learningdisability_psychiatry_cons_new - learningdisability_psychiatry_cons_06
# calculate the percentage difference with 2006
learningdisability_psychiatry_new_pdiff <- (learningdisability_psychiatry_new_ndiff/learningdisability_psychiatry_cons_06) *100

# For Psychotherapy consultants
psychotherapy_cons_06 <- 14.89

psychotherapy_cons_new <- detailed_wte %>%
  filter(job_family == "Medical & dental") %>%
  filter(band_grouping == "Consultant") %>%
  filter(subjob_family == "Psychotherapy") %>%
  summarise(new_sum = sum(new_wte)) %>%
  dplyr::pull(new_sum)

# calculate the difference with 2006
psychotherapy_new_ndiff <- psychotherapy_cons_new - psychotherapy_cons_06
# calculate the percentage difference with 2006
psychotherapy_new_pdiff <- (psychotherapy_new_ndiff/psychotherapy_cons_06) *100

```

__NHS Scotland Workforce__

As at `r new_quarter`,

* NHS Scotland's staffing levels are up by around `r format(round(allwte_new_ndiff, -2), big.mark = ",")` WTE. This is a `r round_half_up(allwte_new_pdiff, 1)`% `r ifelse((allwte_new_pdiff < 0), "decrease", "increase")` (from `r format(round_half_up(allwte_06, 1), big.mark = ",")` WTE at September 2006 to `r format(round_half_up(allwte_new, 1), big.mark = ",")` WTE at `r new_quarter`).

* Total staff paid directly by NHS Scotland as at end `r new_quarter` is now `r format(allhc_new, big.mark = ",")` headcount.

*	Medical & Dental consultant (incl director-level consultant) numbers have `r ifelse ((consultants_new_pdiff < 0),"decreased","increased")` by `r round_half_up(consultants_new_pdiff, 1)`% or `r format(round_half_up(consultants_new_ndiff, 1), big.mark = ",")` WTE since September 2006 to `r format(round_half_up(consultants_wte_new, 1), big.mark = ",")` WTE.

* The Emergency Medicine Consultant workforce `r ifelse ((emergency_cons_06*3 < emergency_cons_new),"has more than tripled","")` from `r round_half_up(emergency_cons_06, 1)` WTE to `r round_half_up(emergency_cons_new, 1)` WTE).

* `r round_half_up(geriatric_new_pdiff, 1)`% `r ifelse((geriatric_new_pdiff < 0), "decrease", "increase")` in Geriatric Medicine Consultants (`r round_half_up(geriatric_cons_06, 1)` to `r round_half_up(geriatric_cons_new, 1)` WTE).

* `r round_half_up(paediatrics_new_pdiff, 1)`% `r ifelse((paediatrics_new_pdiff < 0), "decrease", "increase")` in Paediatrics Specialties Consultants (`r round_half_up(paediatrics_cons_06, 1)` to `r round_half_up(paediatrics_cons_new, 1)` WTE).


__Nursing and Midwifery__

* Since September 2007, there has been `r ifelse((qualnm_new_pdiff < 0), "a decrease", "an increase")` of `r round_half_up(qualnm_new_pdiff, 1)`% qualified nursing and midwifery staff.

* Since September 2007, Qualified Nurses & Midwives have `r ifelse((qualnm_new_pdiff < 0), "decreased", "increased")` by `r round_half_up(qualnm_new_pdiff, 1)`%, or `r format(round_half_up(qualnm_new_ndiff, 1), big.mark = ",")` WTE, from `r format(round_half_up(qual_nmwte_07, 1), big.mark =",")` WTE to `r format(round_half_up(qual_nmwte_new, 1), big.mark =",")` WTE.

__Allied Health Professionals (AHPs)__

*	Under this Government, the number of Allied Health Professionals (excluding paramedics) within NHS Scotland within has   `r ifelse (ahp_new_pdiff < 0,"decreased","increased")` by `r round_half_up(ahp_new_pdiff, 1)`% or `r format(round_half_up(ahp_new_ndiff,1), big.mark =",")` WTE, from `r format(round_half_up(ahpwte_06, 1), big.mark =",")` WTE as at September 2006 to `r format(round_half_up(ahpwte_new, 1), big.mark =",")` WTE as at `r new_quarter`.

__Medical__

* Doctors (excluding dentists) up `r round_half_up(medical_new_pdiff, 1)`% or by `r format(round_half_up(medical_new_ndiff, 1), big.mark = ",")` WTE, (from `r format(round_half_up(medical_wte_06, 1), big.mark =",")` WTE as at September 2006 to `r format(round_half_up(medical_wte_new, 1), big.mark =",")` WTE as at `r new_quarter`).*

* Doctors in Training up `r round_half_up(doctorsintraining_new_pdiff, 1)`% or by `r format(round_half_up(doctorsintraining_new_ndiff, 1), big.mark = ",")` WTE, (from `r format(round_half_up(doctorsintraining_wte_06, 1), big.mark =",")` WTE as at September 2006 to `r format(round_half_up(doctorsintraining_wte_new, 1), big.mark =",")` WTE as at `r new_quarter`).*

*	Medical & Dental consultant (incl director-level consultant) numbers have `r ifelse ((consultants_new_pdiff < 0),"decreased","increased")` by `r round_half_up(consultants_new_pdiff, 1)`% or `r format(round_half_up(consultants_new_ndiff, 1), big.mark = ",")` WTE to `r format(round_half_up(consultants_wte_new, 1), big.mark = ",")` WTE.

* The Emergency Medicine Consultant workforce `r ifelse ((emergency_cons_06*3 < emergency_cons_new),"has more than tripled","")` from `r round_half_up(emergency_cons_06, 1)` WTE to `r round_half_up(emergency_cons_new, 1)` WTE).

* `r round_half_up(geriatric_new_pdiff, 1)`% `r ifelse((geriatric_new_pdiff < 0), "decrease", "increase")` Geriatric Medicine Consultants (`r round_half_up(geriatric_cons_06, 1)` to `r round_half_up(geriatric_cons_new, 1)` WTE).


__Covid-related Specialties__

* Under this Government, the number of Consultants with a specialty of Respiratory Medicine has `r ifelse ((respiratory_new_pdiff < 0 ),"decreased", "increased")` by `r round_half_up(respiratory_new_pdiff, 1)`% (from `r round_half_up(respiratory_cons_06, 1)` WTE as at September 2006 to `r round_half_up(respiratory_cons_new, 1)` WTE as at `r new_quarter`).

* `r round_half_up(infectiousdiseases_new_pdiff, 1)`% `r ifelse ((infectiousdiseases_new_pdiff < 0 ),"decrease", "increase")` in Consultants with a specialty of Infectious diseases (from `r round_half_up(infectiousdiseases_cons_06, 1)` WTE at Sept 2006 to `r round_half_up(infectiousdiseases_cons_new, 1)` WTE at `r new_quarter`).


__Cancer Staffing__

* This Government has seen `r ifelse ((oncology_new_pdiff < 0 ),"a decrease", "an increase")` in the Consultant Oncologist workforce from `r round_half_up(oncology_cons_06, 1)` WTE to `r round_half_up(oncology_cons_new, 1)` WTE - `r round_half_up(oncology_new_pdiff, 1)`% `r ifelse ((oncology_new_pdiff < 0 ),"decrease", "increase")` (September 2006 - `r new_quarter`).

* We've `r ifelse ((radiology_new_pdiff < 0 ),"decreased", "increased")` the Consultant Radiologist workforce from `r round_half_up(radiology_cons_06, 1)` WTE to `r round_half_up(radiology_cons_new, 1)` WTE - a `r round_half_up(radiology_new_pdiff, 1)`% `r ifelse ((radiology_new_pdiff < 0 ),"decrease", "increase")` (September 2006 - `r new_quarter`).

* We've `r ifelse ((radiography_new_pdiff < 0 ),"decreased", "increased")` Radiography staff from `r format(round_half_up(radiography_07, 1), big.mark = ",")` WTE to `r format(round_half_up(radiography_new, 1), big.mark = ",")` WTE since September 2007 - a `r round_half_up(radiography_new_pdiff, 1)`% `r ifelse ((radiography_new_pdiff < 0 ),"decrease", "increase")` (September 2007 - `r new_quarter`).


__Primary Care__

* Paramedics `r ifelse((paramedics_new_pdiff < 0), "down", "up")` `r round_half_up(paramedics_new_pdiff, 1)`% - `r ifelse((paramedics_new_pdiff < 0), "down", "up")` `r round_half_up(paramedics_new_ndiff, 1)` WTE (`r format(round_half_up(paramedics_07, 1), big.mark = ",")` WTE at September 2007 to `r format(round_half_up(paramedics_new, 1), big.mark = ",")` WTE at `r new_quarter`). 


__Mental Health__

* Under this Government, the number of Psychiatric Specialties consultants working in NHS Scotland has `r ifelse ((psychiatry_new_pdiff < 0 ),"decreased", "increased")` by `r round_half_up(psychiatry_new_pdiff, 1)`% (from `r round_half_up(psychiatry_cons_06, 1)` WTE at September 2006 to `r round_half_up(psychiatry_cons_new, 1)` WTE at `r new_quarter`).

* General Psychiatry `r ifelse((genpsychiatry_new_pdiff < 0), "down", "up")` by `r round_half_up(genpsychiatry_new_pdiff, 1)`%
(from `r format(round_half_up(genpsychiatry_cons_06, 1), big.mark = ",")` WTE at September 2006 to `r format(round_half_up(genpsychiatry_cons_new, 1), big.mark = ",")` WTE at `r new_quarter`).

* Child & Adolescent Psychiatry `r ifelse((childpsychiatry_new_pdiff < 0), "down", "up")` by `r round_half_up(childpsychiatry_new_pdiff, 1)`%
(from `r format(round_half_up(childpsychiatry_cons_06, 1), big.mark = ",")` WTE at September 2006 to `r format(round_half_up(childpsychiatry_cons_new, 1), big.mark = ",")` WTE at `r new_quarter`).

* Forensic Psychiatry `r ifelse((forensicpsychiatry_new_pdiff < 0), "down", "up")` by `r round_half_up(forensicpsychiatry_new_pdiff, 1)`%
(from `r format(round_half_up(forensicpsychiatry_cons_06, 1), big.mark = ",")` WTE at September 2006 to `r format(round_half_up(forensicpsychiatry_cons_new, 1), big.mark = ",")` WTE at `r new_quarter`).

* Old age Psychiatry `r ifelse((oldagepsychiatry_new_pdiff < 0), "down", "up")` by `r round_half_up(oldagepsychiatry_new_pdiff, 1)`%
(from `r format(round_half_up(oldagepsychiatry_cons_06, 1), big.mark = ",")` WTE at September 2006 to `r format(round_half_up(oldagepsychiatry_cons_new, 1), big.mark = ",")` WTE at `r new_quarter`).

* Psychiatry of learning disability `r ifelse((learningdisability_psychiatry_new_pdiff < 0), "down", "up")` by `r round_half_up(learningdisability_psychiatry_new_pdiff, 1)`%
(from `r format(round_half_up(learningdisability_psychiatry_cons_06, 1), big.mark = ",")` WTE at September 2006 to `r format(round_half_up(learningdisability_psychiatry_cons_new, 1), big.mark = ",")` WTE at `r new_quarter`).

* Psychotherapy `r ifelse((psychotherapy_new_pdiff < 0), "down", "up")` by `r round_half_up(psychotherapy_new_pdiff, 1)`%
(from `r format(round_half_up(psychotherapy_cons_06, 1), big.mark = ",")` WTE at September 2006 to `r format(round_half_up(psychotherapy_cons_new, 1), big.mark = ",")` WTE at `r new_quarter`).

* Since Sept 2007, Mental Health Nurses have `r ifelse((mhnm_new_pdiff < 0), "down", "up")` by `r round_half_up(mhnm_new_pdiff, 1)`% or by `r format(round_half_up(mhnm_new_ndiff, 1), big.mark = ",")` WTE (from `r format(round_half_up(mh_nmwte_07, 1), big.mark = ",")` WTE as at Sept 2007 to `r format(round_half_up(mh_nmwte_new, 1), big.mark = ",")` WTE at `r new_quarter`).

__NHS Scotland staffing levels have `r ifelse ((allpermanentwte_prevyear_pdiff < 0 ),"decreased", "increased")` by about `r format(round(allpermanentwte_prevyear_ndiff, -1), big.mark = ",")` permanent whole time equivalent in the last year - meaning ten years of growth.__

NHS Scotland staffing levels have `r ifelse ((allpermanentwte_prevyear_pdiff < 0 ),"decreased", "increased")` by `r format(round_half_up(allpermanentwte_prevyear_ndiff, 1), big.mark = ",")` permanent whole time equivalent in the last year - meaning ten years of growth.